---
title: "DUDAS-MADELYNNE-PAEZ_CRESPO-ELSY_JOHANNA-ADA-HOMEWORK-04"
author: "Madelynne Dudas - Elsy Johanna Páez Crespo"
date: "13/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework Assignment 4

<span style="color:blue">***Using Bootstrapping to Estimate Standard Errors and CIs for Linear Models:***</span>

[1] Using the KamilarAndCooperData.csv dataset, run a linear regression looking at log(HomeRange_km2) in relation to log(Body_mass_female_mean) and report your beta coeffiecients (slope and intercept).

```{r}
# Load in dataset
library (curl)
f <- curl("https://raw.githubusercontent.com/difiore/ADA-2019/master/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)

# Run linear regression on log(HomeRange_km2) in relation to log(Body_mass_female_mean)
m <-lm(log(HomeRange_km2)~log(Body_mass_female_mean), data = d)
summary(m)

# Report beta coeffiecients (slope & intercept)
m$coefficients
```

[2] Then, use bootstrapping to sample from your data 1000 times with replacement, each time fitting the same model and calculating the appropriate coefficients. [The size of each sample should be equivalent to the number of observations in your dataset.] This generates a sampling distribution for each beta coefficient. Plot a histogram of these sampling distributions.

```{r}
# Create a loop that samples from data 1000 times with replacement
slopes <- NULL
ints <- NULL
for (i in 1:1000){
  dsamp <- d[sample(nrow(d), 213, replace = TRUE),] #there are 213 observations in the dataset
  m2 <-lm(log(HomeRange_km2)~log(Body_mass_female_mean), data = dsamp) #run linear regressions
  slopes <-append(slopes,m2$coefficients[2]) #add coefficients to a list for each coefficient to generate sampling distributions
  ints <- append(ints,m2$coefficients[1])
}
```

```{r}
# Plot 2 histograms of the sample distributions for slope and intercept
slope_plot <-hist(slopes, col = "white", main = "Sampling Distribution for Slope", 
    xlab = "Slopes", ylab = "Frequency")
int_plot <-hist(ints, col = "white", main = "Sampling Distribution for Intercept", 
    xlab = "Intercepts", ylab = "Frequency")
```

[3] Estimate the standard error for each of your beta coefficients as the standard deviation of the sampling distribution from your bootstrap.

```{r}
# Calculate standard deviation of the sampling distribution for slope and intercept
intsd <- sd(ints)
intsd

slopesd <-sd(slopes)
slopesd
```

[4] Also determine the 95% CI for each of your beta coefficients based on the appropriate quantiles from your sampling distribution.

```{r}
# Calculate 95% CI for slope and intercept
# For intercept
intupper <- mean(ints) + qnorm(0.975, mean = 0, sd = 1) * intsd
intlower <- mean(ints) + qnorm(0.025, mean = 0, sd = 1) * intsd  
intci <- c(intlower, intupper)
intci

# For slope
slopeupper <- mean(slopes) + qnorm(0.975, mean = 0, sd = 1) * slopesd
slopelower <- mean(slopes) + qnorm(0.025, mean = 0, sd = 1) * slopesd  
slopeci <- c(slopelower, slopeupper)
slopeci
```

[5] How does your answer to part [3] compare to the SE estimated from your entire dataset using the formula for standard error implemented in lm()?

```{r}
summary(m)
```
<span style="color:green">**For intercept (see "intsd" (~0.596) compared to 0.67293) The intercept SE is slightly lower from bootstrapping than from that calculated from the entire dataset**</span>

<span style="color:green">**For slope (see "slopesd" (~0.0763) compared to 0.08488) The slope SE is also slightly lower from bootstrapping than from that calculated from the entire dataset**</span>

[6] How does your answer to part [4] compare to the 95% CI estimated from your entire dataset?
```{r}
confint(m)
```

<span style="color:green">**The CIs for both slope and intercept is a little bit narrower in part [4] than for the the entire dataset**</span>

[Extra credit +2]: Write a FUNCTION that takes as its arguments a dataframe ("d"), a linear model ("m", as a character string, e.g., logHR~logBM), a user-defined confidence interval level ("conf.level") with default = 0.95, and a number of bootstrap replicates ("n", with default = 1000). Your function should return a dataframe that includes: beta coefficient names; beta coefficients, standard errors, and upper and lower CI limits for the linear model based on your entire dataset; and mean beta coefficient estimates, SEs, and CI limits for those coefficients based on your bootstrap.

```{r}
lmbootstrap <- function (d, m, conf.level = 0.95, n = 1000, var1, var2) {
m <-lm(var1~var2, data = d) #run linear regression from teh original dataset
slopes <- NULL
ints <- NULL
for (i in 1:1000){
  dsamp <- d[sample(nrow(d), length(var1), replace = TRUE),] #lenght as the lenght of a variable including in the linear model
  m2 <-lm(var1~var2, data = dsamp) #run linear regressions
  slopes <-append(slopes,m2$coefficients[2]) #add coefficients to a list for each coefficient to generate sampling distributions
  ints <- append(ints,m2$coefficients[1])
}
outcome1 <- data.frame(m$coefficients[2], m$coefficients[1], coef(summary(m))[1, "Std. Error"], coef(summary(m))[2, "Std. Error"], confint(m)[1, "97.5 %"], confint(m)[1, "2.5 %"], confint(m)[2, "97.5 %"], confint(m)[2, "2.5 %"])
  outcome2 <- data.frame(mean(slopes), mean(ints), intsd, slopesd, intci[2], intci[1], slopeci[2], slopeci[1])
    colnames(outcome1) <- c("Slope", "Intercept", "SE intercept", "SE slope", "Upper CI (intercept)", "Lower CI (intercept)", "Upper CI (slope)","Lower CI (slope)")
    colnames(outcome2) <- c("Slope", "Intercept", "SE intercept", "SE slope", "Upper CI (intercept)", "Lower CI (intercept)", "Upper CI (slope)","Lower CI (slope)")
    outcomefinal <- rbind(outcome1, outcome2)
    rownames(outcomefinal) <- c("Dataset", "Bootstrap")
    outcomefinal
  return(outcomefinal)
}
```

```{r}
# Using this homework's example:
x <- lmbootstrap(d=d, m=m, var1=log(d$HomeRange_km2), var2=log(d$Body_mass_female_mean))
x
```

[Extra Credit +1]: Graph each beta value from the linear model and its corresponding mean value, lower CI and upper CI from a bootstrap as a function of number of bootstraps from 10 to 200 by 10s.

```{r}
# For slope
slopes <-append(slopes,m2$coefficients[2]) 
v <- seq(from = 10, to = 200, by = 10)
mean <- NULL
intupper <- NULL
intlower <- NULL
for(i in 1:length(v)){
	m <- mean(slopes[1:v[i]])
	mean[i] <- m
	intupper[i] <- m + qnorm(0.975, mean = 0, sd = 1) * slopesd
	intlower[i] <- m + qnorm(0.025, mean = 0, sd = 1) * slopesd 
}

library(ggplot2)
data <- data.frame(mean, intupper, intlower) 
g <- ggplot(data=data, aes(x = v, y = mean))
g <- g + xlab("# of Bootstraps")
g <- g + ylab("Mean of Slopes")
g <- g + geom_point()
g <- g + geom_point(aes(x = v, y = mean), colour = "red")
g <- g + geom_errorbar(aes(ymax= intupper, ymin= intlower))
g
```

```{r}
# For intercept
intercepts <-append(ints,m2$coefficients[2]) 
v1 <- seq(from = 10, to = 200, by = 10)
mean1 <- NULL
intupper1 <- NULL
intlower1 <- NULL
for(i in 1:length(v)){
	m1 <- mean(ints[1:v[i]])
	mean1[i] <- m1
	intupper1[i] <- m1 + qnorm(0.975, mean = 0, sd = 1) * intsd
	intlower1[i] <- m1 + qnorm(0.025, mean = 0, sd = 1) * intsd 
}

library(ggplot2)
data <- data.frame(mean1, intupper1, intlower1)
g <- ggplot(data=data, aes(x = v1, y = mean1))
g <- g + xlab("# of Bootstraps")
g <- g + ylab("Mean of Intercepts")
g <- g + geom_point()
g <- g + geom_point(aes(x = v1, y = mean1), colour = "red")
g <- g + geom_errorbar(aes(ymax= intupper1, ymin= intlower1))
g
```
